---
title: "NHANES 2017-2018 File Prep"
output: html_document
date: '2022-03-18'
---

# Step 1: Read in the data

Here, we will read in the SAS export files from the [NHANES website](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2017){target="_blank"} using the {haven} package.

I have already installed the packages below that I will call with the library() function. To install packages, you simply run install.packages("package name") or go to Tools \> Install Packages \> input package name.

```{r, message=F, warning=F}
library(haven)

demo <- read_xpt("data/DEMO_J.xpt")
bmx <- read_xpt("data/BMX_J.xpt")
pbcd <- read_xpt("data/PBCD_J.xpt")

```

# Step 2: Look at the data

cmd shift m

```{r}
library(tidyverse)
library(skimr)
skim(demo)
skim(bmx)
skim(pbcd)
```

# Step 3: Merge the files

```{r}
bmi_lead <- demo %>% 
  left_join(select(bmx, SEQN, BMXBMI), by = "SEQN") %>% 
  left_join(select(pbcd, SEQN, LBXBPB), by = "SEQN") 
```

# Step 4: Look at the data, part deux

```{r}
skim(bmi_lead)

bmi_lead %>% 
  select(BMXBMI, WTMEC2YR, LBXBPB) %>% 
  skim()

```

## Histograms of BMI and blood lead level

```{r}

hist(bmi_lead$BMXBMI)
hist(bmi_lead$LBXBPB)
```

# Step 5: Recode and format

```{r}
#create our recoded dataset, bmi_lead2
bmi_lead2 <- bmi_lead %>% 
  mutate(#create 4-level bmi categories
         bmicat = case_when(BMXBMI >= 11.5 & BMXBMI <= 24.9 ~ "1",
                            BMXBMI >= 25.0 & BMXBMI <= 29.9 ~ "2",
                            BMXBMI >= 30.0 & BMXBMI <= 39.9 ~ "3",
                            BMXBMI >= 40 ~ "4",
                            TRUE ~ NA_character_),
         #create bmi categories with descriptions
         bmicatf = factor(case_when(bmicat == "1" ~ "Underweight/normal",
                             bmicat == "2" ~ "Overweight",
                             bmicat == "3" ~ "Obese",
                             bmicat == "4" ~ "Severely obese",
                             TRUE ~ NA_character_),
                          levels = c("Underweight/normal", "Overweight", "Obese", "Severely obese"),
                          ordered = T),
         #natural log-transform the blood lead level variable
         ln_lead = log(LBXBPB),
         riagendrf = ifelse(RIAGENDR == 1, "Male", "Female"),
         ridstatrf = ifelse(RIDSTATR == 1, "Interviewed only", "Interviewed and MEC examined"),
         ridreth3f = case_when(RIDRETH3 == 1 ~ "Mexican American",
                               RIDRETH3 == 2 ~ "Other Hispanic",
                               RIDRETH3 == 3 ~ "Non-Hispanic White",
                               RIDRETH3 == 4 ~ "Non-Hispanic Black",
                               RIDRETH3 == 6 ~ "Non-Hispanic Asian",
                               RIDRETH3 == 7 ~ "Other race - including multiracial",
                               TRUE ~ NA_character_),
         ridreth3f = factor(ridreth3f, ordered = F),
         ridreth3f = relevel(ridreth3f, ref = "Non-Hispanic White"),
         targetpop = ifelse(RIAGENDR == 2 & RIDAGEYR >= 20 & !is.na(BMXBMI) & !is.na(LBXBPB) & !is.na(RIDRETH3), 1, 0)
         ) 

```

# Step 6: Descriptive statistics

## Targetpop numbers

```{r}
# how many women in the target pop?
summary(factor(bmi_lead2$targetpop))

#how many women (who would be) in the targetpop who are missing blood metal testing?
bmi_lead2 %>% 
  filter(RIAGENDR == 2 & RIDAGEYR >= 20 & !is.na(BMXBMI) & !is.na(RIDRETH3)) %>% 
  select(LBXBPB) %>% 
  summary()
```

## Distributions of outcome {.tabset .tabset-pills}

### Raw blood lead level

```{r}
hist(bmi_lead2$LBXBPB)
```

### Log-transformed blood lead level

```{r}
hist(bmi_lead2$ln_lead)
```

## Scatterplots {.tabset .tabset-pills}

### Blood lead and BMI

```{r}
bmi_lead2 %>% 
  filter(targetpop == 1) %>% 
  ggplot(aes(x = BMXBMI, y = LBXBPB)) +
  geom_point()
```

### Log-transformed blood lead and BMI

```{r}
bmi_lead2 %>% 
  filter(targetpop == 1) %>% 
  ggplot(aes(x = BMXBMI, y = ln_lead)) +
  geom_point()
```

# Step 7: Saving an R dataset for later use

```{r}
write_rds(bmi_lead2, "bmi_lead2.rds")
```
