---
title: "NHANES 2017-2018 Linear Regression"
output: html_document
date: '2022-03-18'
---

In our nhanes_data-prep file, we created our analytic dataset to use for our linear regression. We will read in the dataframe we made in that file here and then run unadjusted and adjusted linear regression models.

n = 2569

```{r}
library(tidyverse)
library(survey)
library(jtools)
```

# Read in dataset
```{r}
bmi_lead2 <- read_rds("bmi_lead2.rds")
```

# Descriptive analysis for Table 1

## Make survey design object
```{r, message=FALSE, warning=FALSE}
# Make the survey design object
nh_svy <- svydesign(data = bmi_lead2, strata = ~SDMVSTRA, id = ~SDMVPSU, weights = ~WTMEC2YR, nest = T)

# For working with a subset of the population, we can do the following. 
target_svy <- subset(nh_svy, targetpop == 1)
```

## Get Table 1 numbers

```{r}
bmi_lead2 %>% 
  filter(targetpop == 1) %>% 
  group_by(bmicatf) %>% 
  summarise(n = n())
```

```{r}
t1_1 <- svyby(~RIDAGEYR, ~bmicatf, design = target_svy, svymean)

ftable(t1_1) %>% round(2)

t1_1 %>% 
  mutate(RIDAGEYR = round(RIDAGEYR, 1),
         se = round(se, 2))
```

```{r}
t1_2 <- svyby(~ridreth3f, ~bmicatf, design = target_svy, svymean)

t1_2 %>% 
  # here I'm just getting my output in the same order as Kate's table:
  select(contains("Mexican American"), contains("Other Hispanic"), contains("Non-Hispanic Black"), contains("Non-Hispanic White"), contains("Non-Hispanic Asian"), contains("Other")) %>% 
  mutate(across(.cols = starts_with("ridreth")), .*100) %>% 
  round(1)

```


# Linear regression ignoring complex survey design

```{r}
# Make a df for the target population
targetpop_df <- bmi_lead2 %>% 
  filter(targetpop == 1)

# define the model object using linear model, ignoring complex survey design
model1 <- lm(LBXBPB ~ BMXBMI, data = targetpop_df)
 
# Plot residuals vs fitted, Q-Q, Scale-location, residuals vs leverage
plot(model1)

# A clean way to look at your results 
broom::tidy(model1, conf.int = T) 

# Another way to look at your results
jtools::summ(model1, confint = T)
```


# Linear regression with complex survey design
```{r}
model_svy_1 <- svyglm(LBXBPB~BMXBMI, design=target_svy)

summary(model_svy_1)

plot(svyglm(LBXBPB~BMXBMI, design=target_svy))

broom::tidy(model_svy_1)

jtools::summ(model_svy_1, confint = T)
```

## Simple

```{r}
model_svy_2 <- svyglm(ln_lead~BMXBMI, design=target_svy)
summary(model_svy_2)
plot(svyglm(ln_lead~BMXBMI, design=target_svy))
broom::glance(model_svy_2)
```

## Multivariable
```{r}
summary(svyglm(LBXBPB ~ BMXBMI + RIDAGEYR + ridreth3f, design=target_svy))
plot(svyglm(ln_lead~BMXBMI + RIDAGEYR + ridreth3f, design=target_svy))
```

## Stratified by race/ethnicity
```{r}

```

## BMI * Age interaction
```{r}

```

